---
title: "Get Started"
author: "Shu Fai Cheung & Mark Hok Chio Lai"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
csl: apa.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width  =  6,
  fig.height =  6,
  fig.align = "center",
  fig.path = ""
)
```

# Introduction

This article illustrates how to use
[`semlrtp`](https://sfcheung.github.io/semlrtp/) to
compute LRT *p*-values for selected free parameters in
a model fitted by `lavaan`.

These are the packages needed for this illustration:

```{r}
library(lavaan)
library(semlrtp)
```

# LRT *p*-Values

What we called the LRT *p*-value of a parameter is
simply the *p*-value of the likelihood
ratio test (LRT, a.k.a. $\chi^2$ difference test)
comparing the fitted model with the model with
this parameter fixed to zero. It is not new but
we are not aware of packages using it as the default
*p*-value in testing model parameters.
The package
[`semlrtp`](https://sfcheung.github.io/semlrtp/)
is developed to facilitate the use of
this *p*-value.

# Basic Workflow

## Data

This is the sample dataset from the package,
with 16 variables and a group variable:

```{r}
data(data_sem16)
print(head(data_sem16), digits = 2)
```

## Model

This is a model to be fitted, with
four latent factors, and a structural
model for the factors:

```{r}
mod <-
"
f1 =~ x1 + x2 + x3
f2 =~ x5 + x6 + x7
f3 =~ x9 + x10 + x11
f4 =~ x13 + x14 + x15
f3 ~ f1 + f2
f4 ~ f3
"
```

We first fit a model to the whole sample:

```{r}
fit <- sem(mod,
           data_sem16)
```

## LRT *p*-Values For Selected Parameters

If we use the default settings,
we can compute the LRT *p*-values just
by calling `lrtp()` on the `lavann`
output.

By default, LRT *p*-values will be computed
only for regression paths
(`"~"`) and covariances (`"~~"`),
excluding variances and error covariances.

```{r}
fit_lrtp <- lrtp(fit)
```

By default, the output will be printed
in `lavaan` style, and only parameters
with LRT *p*-values are printed:

```{r}
fit_lrtp
```

The column `Chisq` shows the $\chi^2$
difference of the likelihood ratio test
when a parameter is fixed to zero.

The column `LRTp` shows the LRT *p*-value.

## How LRT *p*-Values Are Computed

It can be verified that the LRT *p*-value
of a parameter is the likelihood ratio
test (LRT) *p*-value (a.k.a. the $\chi^2$
difference test) when this parameter is
fitted to zero.

For example, we fix the path from `f2`
to `f3` to zero and then do an LR test:

```{r}
mod_f3_f2 <-
"
f1 =~ x1 + x2 + x3
f2 =~ x5 + x6 + x7
f3 =~ x9 + x10 + x11
f4 =~ x13 + x14 + x15
f3 ~ f1 + 0*f2
f4 ~ f3
"
fit_f3_f2 <- sem(mod_f3_f2,
                 data_sem16)
lavTestLRT(fit_f3_f2,
           fit)
```

```{r echo = FALSE}
est <- parameterEstimates(fit)
f3_f2_wald_p <- est[14, "pvalue"]
f3_f2_lrt_p <- fit_lrtp[14, "LRTp"]
```

Unlike the original *p*-value of
this path,
`r formatC(f3_f2_wald_p, digits = 3, format = "f")`,
the LRT *p*-value is
`r formatC(f3_f2_lrt_p, digits = 3, format = "f")`,
suggesting that the path from `f2` to `f2`
is significant.

# Why LRT *p*-Value

The example above illustrates the importance
of the LRT *p*-value. The usual *p*-values
in `lavaan` (and many other SEM
programs) are Wald-based *p*-values.
The Wald-based p-value is an approximation
of the
LRT *p*-value when a parameter is fixed
to zero. It is an approximation and
so can be different from the
LRT *p*-value. Moreover, they may also
depend on
the parameterization [@gonzalez_testing_2001].

For example, we can fit the same model
by changing the indicators being fixed to 1.

```{r}
mod2 <-
"
f1 =~ x2 + x1 + x3
f2 =~ x7 + x5 + x6
f3 =~ x11 + x9 + x10
f4 =~ x14 + x13 + x15
f3 ~ f1 + f2
f4 ~ f3
"
fit2 <- sem(mod2,
            data_sem16)
```

This model and the previous have exactly
identical model fit, as expected:

```{r}
fitMeasures(fit, c("chisq", "df"))
fitMeasures(fit2, c("chisq", "df"))
```

This is the parameter estimates with
Wald *p*-values (only those for
the regression paths are displayed)

```{r echo = FALSE}
est2 <- parameterEstimates(fit2, output = "text")
tmp <- capture.output(est2)
f3_f2_wald_p2 <- est2[14, "pvalue"]
```

```r
parameterEstimates(fit2,
                   output = "text")
```

```{r echo = FALSE}
cat(tmp[21:27], sep = "\n")
```

The Wald *p*-value is
`r formatC(f3_f2_wald_p2, digits = 3, format = "f")`,
even larger than
`r formatC(f3_f2_wald_p, digits = 3, format = "f")`
in the original model.

However, the LRT *p*-values are the same:

```{r}
fit2_lrtp <- lrtp(fit2)
fit2_lrtp
```

It is because LRT *p*-value is invariant to
parameterization.

# Limitations

Because LRT *p*-value are computed by
fixing a parameter to zero, there are
parameter for which the LRT *p*-value
cannot be computed. For example,
suppose we request LRT *p*-values for
factor loadings using the argument
`op`:

```{r}
fit_lrtp_loadings <- lrtp(fit,
                          op = "=~")
fit_lrtp_loadings
```

As shown above, LRT *p*-values are not
computed for indicators with loadings
fixed to zero.

# Further Information

Please refer to the help page of
`lrtp()` for other arguments, and
the print method of `lrtp()` output
(`print.lrtp()`) for options
in printing.

# References
